---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

## dplyrodbc

```{r,eval=FALSE}
devtools::install_github("mdsumner/RODBCDBI",   ref = "mike")
```


## Example

ODBC has an installed File DSN called "CountriesAccess". This is a .mdb file created using 32-bit Windows (effectively) and configured with the Data Sources application from "C:\Windows\SysWOW64\odbcad32.exe". 

Minimal test for RODBC. 

```{r}
library(RODBC)  ## remember this is 32-bit R
con <- odbcConnect("CountriesAccess1")
## some columns explode the app, presumably "Geom (I)"
sqlQuery(con, "SELECT TOP 10 * FROM [Countries Table]")
close(con)
```

Methods for us. 

```{r}

src_access <- function(dbname = NULL, host = NULL, port = NULL, user = NULL,
                         password = NULL, ...) {

  con <-    dbConnect(RODBCDBI::ODBC(), dsn=dbname)

  src_sql("access", con)
}



src_desc.src_access <- function(con) {
  info <- dbGetInfo(con$con)
  host <- if (info$host == "") "localhost" else info$host

  paste0("Access ", info$serverVersion, " [", info$user, "@",
    host, ":", info$port, "/", info$dbname, "]")
}


db_list_tables.src_access <- function(con) {
  dbListTables(con$con)
}

db_has_table.src_access <- function(con, table) {
  dbExistsTable(con$con, table)
}

tbl.src_access <- function(src, from, ...) {
  tbl_sql("access", src = src, from = from, ...)
}

src_translate_env.src_access <- dplyr:::src_translate_env.NULL

sql_squote <- 
function (x) 
{
 # y <- gsub(quote, paste0(quote, quote), x, fixed = TRUE)
  y <- paste0("[", x, "]")
  y[is.na(x)] <- "NULL"
  names(y) <- names(x)
  y
}

sql_escape_string.ODBCConnection <- function(con, x) {
   sql_squote(x)
}

sql_escape_ident.ODBCConnection<- function(con, x) {
  sql_squote(x)
}
```


Try for real with our Access source. Here we see the engine generate a SQL query with all of the column names and table name escaped correctly, but with LIMIT 10. How to modify this to use TOP 10? 

```{r}
library(DBI)
library(RODBCDBI)
library(dplyr)
x <- src_access("CountriesAccess1")

## a debug statement to print(res@sql) in the dbFetch/fetch methods shows the problem with LIMIT
e <- try(tbl(x, 'Countries Table'))

```

> From: Countries Table [3,063 x 14]
> 
> <SQL> SELECT [ID], [TypeI], [BranchesI], [CoordinatesI], [XI], [YI], [LongitudeI], [LatitudeI], [LengthI], [AreaI], 
> [BearingI], [SelectionMaskI], [SelectionI], [VersionI]
> FROM [Countries Table]
> LIMIT 10









